---
title: "dataClean"
author: "Yiwen Peng"
date: "2024-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyverse, lubridate, patchwork)
```


```{r}
# data1 <- readxl::read_xls("E:/Ariel/adelaide/research/data/2000.xls")
# data2 <- readxl::read_xls("E:/Ariel/adelaide/research/data/2001.xls")
# # check two table columns' type
# str(data1)
# str(data2)

```
## data cleaning

2000 没有betting odds, 所以不合并选择这一年的数据
合并其它年份，(每个column的type是否一致)

```{r}
#check missing value
#row number
# total <- nrow(data2)
# summary(data2)
```

missing value
选择betting odds 的时候， GBW             GBL             IWW             IWL  可以选择来计算elo，因为他们的缺失值468 and 469, missing value 最少，最适合直接drop.

2. calculate each player's betting odds mean and replace missing value


```{r}
read_xls_file <- function(file_path) {
  df <- readxl::read_xls(file_path)
  # Convert WRank column to double
  
  # Check if 'WPts' column exists
  if("WPts" %in% names(df)) {
    # Convert specified columns to integers
    df <- df %>%
      mutate(WRank = as.integer(WRank),
             LRank = as.integer(LRank),
             WPts = as.integer(WPts),
             LPts = as.integer(LPts))
  } else {
    # If 'WPts' column doesn't exist, set it to NA
    df <- df %>%
      mutate(WRank = as.integer(WRank),
             LRank = as.integer(LRank),
             WPts = NA_integer_,
             LPts = NA_integer_)
  }
}


  read_xlsx_file <- function(file_path) {
  df <- readxl::read_xlsx(file_path)
  # Convert WRank column to double
  
  # Check if 'WPts' column exists
  if("WPts" %in% names(df)) {
    # Convert specified columns to integers
    df <- df %>%
      mutate(WRank = as.integer(WRank),
             LRank = as.integer(LRank),
             WPts = as.integer(WPts),
             LPts = as.integer(LPts))
  } else {
    # If 'WPts' column doesn't exist, set it to NA
    df <- df %>%
      mutate(WRank = as.integer(WRank),
             LRank = as.integer(LRank),
             WPts = NA_integer_,
             LPts = NA_integer_)
  }
}
```


### 合并其它年份
2000 to 2012

2000的column 和其他表格的不一样，所以drop 2000
```{r}
# Import data from 2000 - 2012
files1 <- str_glue("E:/Ariel/adelaide/research/data/{2000:2012}.xls")
raw_matches1 <- files1 %>%
  map_dfr(function(x)  read_xls_file(x))
```
20
```{r}
# Import data from 2013 - 2024
files2 <- str_glue("E:/Ariel/adelaide/research/data/{2013:2024}.xlsx")
raw_matches2 <- files2 %>%
  map_dfr(function(x)  read_xlsx_file(x))

```

```{r}
print(nrow(raw_matches1))
print(nrow(raw_matches2))
print(36541 +27805)
```

combine two tibble 

```{r}
raw_matches <- bind_rows(raw_matches1,raw_matches2,.id = "source")
```

```{r}
raw_matches
```



```{r}
str(raw_matches)
skimr::skim_without_charts(raw_matches)
```

图三可以看出，有很多列的complete rate  都很小，只有0.0几。complete rate 不到0.5的都可以drop ,缺失值太多了

 LRank     : chr [1:8678] "101" "442" "485" "69" ...
 change this to number type
```{r}
raw_matches$LRank <- as.numeric(raw_matches$LRank)
typeof(raw_matches$LRank)
```


```{r}
str(raw_matches)
```

extract player information

```{r}
# c(unique(raw_matches$Winner),unique(raw_matches$Loser))
# 
# unique(c(unique(raw_matches$Winner),unique(raw_matches$Loser)))

p1 <- unique(raw_matches$Winner)
p2 <- unique(raw_matches$Loser)
p3 <- unique(c(p1,p2))
print(length(p3))
print(length(unique(raw_matches$Winner)))
print(length(unique(raw_matches$Loser)))
print(length(c(unique(raw_matches$Winner),unique(raw_matches$Loser))))
print(length(unique(c(unique(raw_matches$Winner),unique(raw_matches$Loser)))))
```
```{r}
players_data <- tibble(player = p3)

```

```{r}
players_data <- players_data %>% mutate(
  Elo_score = 1500,
  t_times = 1
)

players_data
```



```{r}
saveRDS(players_data, file = "E:/Ariel/adelaide/research/R/data/players.rds")
```

```{r}
c(unique(raw_matches$Winner),unique(raw_matches$Loser))

unique(c(unique(raw_matches$Winner),unique(raw_matches$Loser)))

```



